\ifx\mainfile\undefined
\input{chapter_head}
\setcounter{chapter}{7} % Set to n-1!
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cleardoublepage
\chapter{Further Applications of Effective SNR}
\label{chap:applications}

In the previous chapter, I showed that my Effective SNR model provides good performance when selecting rates for 802.11a/g and 802.11n links. To further illustrate the model's flexibility and power, this chapter presents my exploration of Effective SNR to various problems of wireless links and wireless networks.

\begin{table}[htp]
	\centering
	\begin{tabular}{lc}
	\toprule
		\textbf{Application of Effective SNR} & \textbf{Described in} \\
	\midrule
		Bitrate/MCS selection & \chapref{chap:rate}, originally in \cite{Halperin_ESNR}\\
		Channel width selection & \chapref{chap:rate}, originally in \cite{Halperin_ESNR}\\
		Antenna selection & \chapref{chap:rate}, originally in \cite{Halperin_ESNR}\\
		Power control & \chapref{chap:rate}, originally in \cite{Halperin_ESNR}\\
		Channel selection & \secref{sec:esnr_chansel}\\
		AP selection & \secref{sec:esnr_apsel}\\
		Path selection/BSS selection in WDS & \secref{sec:esnr_pathsel}\\
		Interference planning \\
		Partial packet recovery/FEC & Bhartia et al.~\cite{Bhartia_FreqDiv}\\
		Beamforming \\
		Multicast rate selection \\
	\bottomrule
	\end{tabular}
	\caption[A variety of applications of Effective SNR]{\label{tab:esnr_uses}A variety of applications of Effective SNR\@.}
\end{table}

\tabref{tab:esnr_uses} shows a list of several potential applications of Effective SNR, as those described in \chapref{chap:problem}. These range from optimizing various parameters of a single Wi-Fi link, such as the MCS or antenna set used, to coordinating many 802.11 nodes in a dense wireless network. Additionally, I identify applications that can be implemented by looking at other aspects of the Channel State Information below in \tabref{tab:csi_uses}. These provide useful primitives that can enable systems to adapt behavior based on the location and movement of the user. Combined, I believe these form the critical building blocks for configuring dense 802.11 networks.

\begin{table}[htp]
	\centering
	\begin{tabular}{lc}
	\toprule
		\textbf{Application of CSI} & \textbf{Described in} \\
	\midrule
		Mobility classification & \secref{sec:esnr_mobility}\\
		Indoor localization & Sen et al.~\cite{Sen_PinLoc,Sen_SpinLoc} \\ 
	\bottomrule
	\end{tabular}
	\caption[A variety of applications of Channel State Information]{\label{tab:csi_uses}A variety of applications of Channel State Information.}
\end{table}

In this chapter, I explore CSI and Effective SNR approaches to implementing four of the above applications. The first three are are selection algorithms for access points, operating channels, and multi-hop paths through networks that represent the main decisions (beyond rate) that need to be made in flexible topologies such as Wi-Fi Direct networks. The last application is to classify whether a device is mobile, to switch between algorithms designed for static vs mobile links, or trigger a search for a new, better operating point when the device begins to move.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{esnr_eval_dataset}
\input{esnr_eval_apsel}
\input{esnr_eval_chansel}
\input{esnr_eval_pathsel}
\input{esnr_eval_mobility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
In the previous two chapters, I showed that my Effective SNR model can predict packet delivery for MIMO and OFDM networks, and that these predications can be used to accurately select rates for 802.11a/g and 802.11n links. In this chapter, I focused on problems that go beyond simple rate selection for a link. In particular, I evaluated three algorithms using Effective SNR and one algorithm using Channel State Information that both illustrate the flexibility of my model and also flesh out the application space of Wi-Fi network configurations.

I showed that Effective SNR can be used by a client to choose between different access points when connecting to the network, by a pair of devices initiating a peer-to-peer link to choose the best operating channel, and by a client in conjunction with a set of relays to choose a good path to a destination. These three applications are key components of network-level configuration algorithms that are increasingly relevant in the evolving Wi-Fi space; my Effective SNR-based algorithms make them practical. Effective SNR provides strong performance improvements over the Packet SNR for the former two applications. Interestingly, I found that although Effective SNR works to inform multi-hop path selection in the network, it does not perform significantly better than Packet SNR because of the 802.11n spatial diversity techniques that improve the accuracy of Packet SNR.

Finally, inspired by recent work that uses information from onboard sensors about whether a device is mobile to improve link- and network-level performance, I designed and evaluated a pure CSI-based mobility classifier. This classifier can be used to complement the prior applications in a variety of ways, such as informing the nodes in the network when it might be worth looking for a new operating point.

This chapter concludes the evaluation of my Effective SNR model for future wireless networks and my exploration of its power. The applications I have presented here capture the inherent flexibility of this model, and it can be used in many more applications. In the next chapter, I discuss where my work fits in the space of related research on wireless networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifx\mainfile\undefined
\input{chapter_tail}
\fi