\cleardoublepage
\acknowledgments{% \vskip2pc
  % {\narrower\noindent
  Writing acknowledgements is always tricky business, because one invariably forgets someone. To circumvent this pitfall, I chose someone very important and left them out deliberately. You know who you are.

  I feel extremely privileged to have worked as part of an amazing community of people during my graduate studies. First and foremost, I am immensely grateful to David Wetherall and Tom Anderson, my amazing and inspiring advisors. They have guided me through the research process, taught me how to frame problems and analyze research, and have supported me no matter what over the past six years. Beyond this, David's unfailingly positive attitude and his excellent advice renewed my spirits when I was low on energy; I always walked out of our meetings with a newfound vigor and enthusiasm for my work. And Tom's ability to get to the heart of an issue and distill the key points has helped me focus my work. These are only a few ways in which I have benefited tremendously from both relationships.
  
  I have enjoyed collaborating with many other professors, students, and researchers over the years. In particular, Wenjun Hu and Anmol Sheth were great mentors and key contributors to my thesis work. I was fortunate to work with Victor Bahl, Jitu Padhye, Srikanth Kandula, Yoshi Kohno, Arvind Krishnamurthy, Ben Greenstein, Kevin Fu, Bill Maisel, Tom Heydt-Benjamin, George Nychis, Ben Ransford, Vincent Liu, Josie Ammer, Shane Clark, Benessa Defend, Dongsu Han, Tom Kenney, Will Morgan, Eldad Perahia, Srini Seshan, Robert Stacey, and Peter Steenkiste over the years, and I wish we could have worked together more.
  
  I thank all the inhabitants of the networking lab over the years for making life fun inside and outside the lab, creating a work environment I wanted to visit seven days a week (and often did). I also thank all my friends and colleagues at UW CSE---the kind of place where those words are synonymous---for creating an atmosphere of learning, respect, and friendship where it is easy to succeed. I would especially like to thank Neva Cherniavsky, Tomas Isdal, and Ratul Mahajan for always being there with advice and emotional support and serving as great examples of success.
  
  None of us, especially me, would have gotten anything done in graduate school were it not for the tireless and heroic efforts of Melody Kadenko, our program manager, and Lindsay Michimoto, our graduate program advisor. They kept our \emph{i}s dotted and our \emph{t}s crossed, and generally made sure we stayed in line. They gracefully and graciously handled every request no matter how last-minute (or late). I also appreciate the support of Karla Danson and Julie Svendsen.
  
  I would like to thank all my roommates, Saleema Amershi, Sal Guarnieri, Alex Jaffe, Nodira Khoussainova, Ian McDonald, and Dustin Shilling. Living with them was always a pleasure and was especially important in making it possible to get through deadlines. Along with Neva, Tomas, Didi, Mike, John, Kayur, and many more, they always reminded that it was important to play as well as work.
  
  My graduate work was supported by the National Science Foundation, the UW Clairmont L. Egtvedt Fellowship, an Intel Foundation Ph.D. Fellowship, and through internships at Intel Labs Seattle and Microsoft Research.
  
  Most importantly, I'd like to thank my family. They have been there for me through every stage of my life, and their encouragement and support have made all the difference.
  % \par}
}