\ifx\mainfile\undefined
\input{chapter_head}
\setcounter{chapter}{7} % Set to n-1!
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cleardoublepage
\chapter{Further Applications of Effective SNR}
\label{chap:applications}

The detailed evaluation presented in the previous chapter showed that my Effective SNR model provides good performance when selecting rates for 802.11a/g and 802.11n links, one important application in wireless link configuration. In this chapter, I present an evaluation of the Effective SNR for four wireless configuration problems to further illustrate the broad applicability of my work.

I chose four problems that represent the key applications needed to configure a future wireless network. The first three are selection algorithms for access points, operating channels, and multi-hop paths through networks. These represent the main decisions (beyond rate) that need to be made in flexible topologies such as Wi-Fi Direct networks. The last application is to classify whether a device is mobile, in order to switch between algorithms designed for static vs. mobile links, or to trigger a search for a new, better operating point when the device begins to move. Together these are the building blocks of an automatically-reconfiguring network that provides good performance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{esnr_eval_dataset}
\input{esnr_eval_apsel}
\input{esnr_eval_chansel}
\input{esnr_eval_pathsel}
\input{esnr_eval_mobility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}
In this chapter, I evaluated three algorithms using Effective SNR and one algorithm using Channel State Information to illustrate the flexibility of my model and also flesh out the space of Wi-Fi network configurations. My Effective SNR-based algorithms offer good performance for these applications and make them practical.

Effective SNR provides strong performance improvements over the Packet SNR for the access point and channel selection applications. Interestingly, I found that although Effective SNR can inform multi-hop path selection in the network, it does not perform significantly better than Packet SNR because of the 802.11n spatial diversity techniques that improve the accuracy of Packet SNR.

Finally, inspired by recent work that uses information from on-board sensors to improve link- and network-level performance, I designed and evaluated a pure CSI-based mobility classifier. This classifier can be used to complement the prior applications in a variety of ways, such as informing the nodes in the network when it might be worth looking for a new operating point.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifx\mainfile\undefined
\input{chapter_tail}
\fi